---
title: "Clustering"
author: "Connor Flynn"
date: "5/21/2023"
output: html_document
---


Goal of this Rmd: Cluster countries based on ND-GAIN Data

Steps:
- Reshape data
- Impute missing Data
- Determine optimal # of clusters
- Cluster data
- Analyze results

Load in Libraries

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(dplyr)
library(cluster)
library(factoextra)
library(ClustImpute)
library(caret)
library(Hmisc)
#library(estlibrary)

options(scipen = 99)
options(ggrepel.max.overlaps = Inf)
set.seed(3)
```



Read in Data

[ND-GAIN Country Index](https://gain.nd.edu/our-work/country-index/)

The ND-GAIN Country Index summarizes a country's vulnerability to climate change and other global challenges in combination with its readiness to improve resilience. It aims to help governments, businesses and communities better prioritize investments for a more efficient response to the immediate global challenges ahead.


```{r}
nd_gain_data <- read_csv(here("SIDS_Data_Platform_ND-GAIN/data/nd-gain-joined.csv"))
```


Round values in the value column to 2 digits after the decimal

```{r}
nd_gain_data$value <- round(nd_gain_data$value, digits = 2)
```


**Challenge**

The dataset is slightly flawed because if a value is not measured one year, it simply gets carried over from the previous year

So simply think of filtering for the year 2020 as using the most recent year of available data (warning: this can lead to 20 year old data being visualized)


```{r}
nd_gain_data <- nd_gain_data %>% 
  filter(year == 2020)
```

Shift dataframe so that each value in the indicator column becomes it's own column 
 - This gives us 190 rows (one country per row) with 48 indicators

```{r}
nd_gain_data <- nd_gain_data %>% 
  pivot_wider(names_from = indicator, 
              values_from = value)
  
```

Remove non-numeric columns

```{r}
nd_gain_data <- nd_gain_data %>% 
  select(-ISO3, -year, -sids_context, -population, -innovation)
```

Make country names the row name 

```{r}
nd_gain_data <- nd_gain_data %>% 
  remove_rownames %>% 
  column_to_rownames(var="Name")
```

Only keep columns where at least 80% of the data is present

```{r}
nd_gain_data_less_na <- nd_gain_data[, which(colMeans(!is.na(nd_gain_data)) > 0.8)]

```



Impute using ClustImpute

* For nr_cluster = x, we are determining the number of clusters that we believe there are in the data
  - This number can be changed and has significant influence on the output data

```{r}
res <- ClustImpute(nd_gain_data_less_na, nr_cluster = 2)
```


Go into the ClustImpute object and retrieve the imputed data frame

```{r}

nd_gain_data_Clust_Impute <- res$complete_data

```


Normalize the Data

```{r}
process <- preProcess(as.data.frame(nd_gain_data_Clust_Impute), method=c("range"))

normalized_nd_gain <- predict(process, as.data.frame(nd_gain_data_Clust_Impute))
```




Histograms

```{r}
#hist.data.frame(normalized_nd_gain)

```


This visualization helps us to determine the number of clusters that should be used in the kmeans() function

```{r}
fviz_nbclust(normalized_nd_gain, kmeans, method = "silhouette")
```



Cluster the newly imputed data
```{r}
set.seed(3)
kmeans2 <- kmeans(normalized_nd_gain, center = 2, nstart = 100)
```

Visualize clusters
```{r}
fviz_cluster(kmeans2, data = normalized_nd_gain,
geom = c("text","point"),
             repel = TRUE,            # Avoid label overlapping
             show.clust.cent = TRUE, # Show cluster centers
             palette = "lancet",         # Color palette see ?ggpubr::ggpar
             ggtheme = theme_bw()+
              theme(axis.text.x = element_text( hjust = 0.5, vjust = 0.5, size=2),
        axis.title.x =element_text(size=20),
        axis.text.y = element_text(hjust = 0.5, vjust = 0.5, size=2),
        axis.title.y =element_text(size=2),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black"),
        strip.text = element_text(size=2)))
```
View Clusters as Dataframe

```{r}
nd_gain_data_Clust_Impute_clusters <- as.data.frame(kmeans2$cluster)
```

Put countries back into column Name

```{r}
nd_gain_data_Clust_Impute_clusters <- rownames_to_column(nd_gain_data_Clust_Impute_clusters, "Name")
```

Give the cluster column the name cluster2

```{r}
names(nd_gain_data_Clust_Impute_clusters)[names(nd_gain_data_Clust_Impute_clusters) == "kmeans2$cluster"] <- "cluster2"
```

Write this as it's own csv file so it can be joined with the main ND-GAIN data

```{r}
 write_csv(nd_gain_data_Clust_Impute_clusters, here("SIDS_Data_Platform_ND-GAIN/data/two_cluster_data.csv"))
```

