---
title: "ND- GAIN Data Correlation Matrix"
author: "Connor"
date: "5/19/2023"
output: html_document
---

Load Packages and Libraries

```{r setup, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(tidyverse)
library(ggcorrplot)
library(plotly)
library(naniar)
```

Read in the data


```{r}
dataset <- read_csv(here("SIDS_Data_Platform_ND-GAIN/data/nd-gain-joined.csv"))
```


Round decimals in the value column


```{r}
dataset$value <- round(dataset$value, digits = 2)
```




Select for Year 2020


**Challenge**


The dataset is slightly flawed because if a value is not measured one year, it simply gets carried over from the previous year

```{r}
dataset_2020 <- dataset %>% 
  filter(year == 2020)

dataset_2015 <- dataset %>% 
  filter(year == 2015)

sum(is.na(dataset_2020$value))
sum(is.na(dataset_2015$value))



```

So simply think of this as using the most recent year of available data (warning: this can lead to 20 year old data being visualized)


```{r}
dataset_most_recent <- dataset %>% 
  filter(year == 2020)
```

Pivot Wider
We do this so that instead of having an indicator column with each indicator as a value in that column, we have each indocator as it's own column in the dataset. 
- This gives us 190 rows (one country per row) with 48 indicators

```{r}
dataset_most_recent_wider <- dataset_most_recent %>% 
  pivot_wider(names_from = indicator, 
              values_from = value)
  
```


To create our correlation matrix, we only want columns with numeric values to be in our datframe so here, we take out non-numeric columns


```{r}
dataset_most_recent_wider_numeric <- dataset_most_recent_wider %>% 
  select(-ISO3, -year, -Name, -sids_context)
```


Here, we create the correlation matrix

```{r}
cor_matrix <- cor(dataset_most_recent_wider_numeric[complete.cases(dataset_most_recent_wider_numeric), ], use = "pairwise.complete.obs")
```

Plot the Correlation matrix

```{r}
corrplot <- ggcorrplot(cor_matrix, type = "lower", outline.color = "black") +
  theme(axis.text.x = element_text(size = 3),
        axis.text.y = element_text(size = 3)) 


corrplot
```

How to Interpret:

Here we have all of our indicators on the x an y axis. With this visualization, the relationships between each variable are described with color. 

- Looking at the legend, we see a range from dark red (inidcating a strong positive correlation) to dark purple (indicating a strong negative correlation)
  - Positive Correlation -> as one indicator (variable) increases, so does the other / as one indicator (variable) decreases, so does the other 
  - Negative Correleation -> as one indicator (variable) increases, the other indicator (variable) decreases / as one indicator (variable) decreases, the other indicator (variable) increases 
  
*Always test what you see in a correlation matrix by putting the two correlated variables into a scatter plot*


Make the Correlation Matrix Plot Interactive with ggplotly

```{r}
ggplotly(corrplot)
```


Filter for SIDS

```{r}
dataset_most_recent_wider_sids <- dataset_most_recent_wider %>% 
  filter(sids_context == "sids")
```


Select only the numeric columns

```{r}
dataset_most_recent_wider_sids_numeric <- dataset_most_recent_wider_sids %>% 
  select(-ISO3, -Name, -year, -sids_context, -innovation)
```


```{r}
cor_matrix_sids <- cor(dataset_most_recent_wider_sids_numeric, use = "pairwise.complete.obs")
```



```{r}
corrplot_sids <- ggcorrplot(cor_matrix_sids, type = "lower", outline.color = "white") +
  theme(axis.text.x = element_text(size = 3),
        axis.text.y = element_text(size = 3))

corrplot_sids
```

```{r}
ggplotly(corrplot_sids)
```

