---
title: "ND GAIN Data Exploration"
author: "Connor"
date: "6/13/2023"
output: html_document
---

Load Libraries

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)# allows us to cut out long file paths (ex. "users/connor/dowloads/etc")
library(tidyverse) # allows us to do lots of fundamental data science tasks (ex. read_csv)
library(ggcorrplot) # allows us to make correlation plots
library(plotly) # allows us to make ggplot objects interactive
library(ggpubr) # stat_cor for correlation coefficients
```

Read in Data

[ND-GAIN Country Index](https://gain.nd.edu/our-work/country-index/)

The ND-GAIN Country Index summarizes a country's vulnerability to climate change and other global challenges in combination with its readiness to improve resilience. It aims to help governments, businesses and communities better prioritize investments for a more efficient response to the immediate global challenges ahead.

```{r} 
nd_gain_data <- read_csv(here("SIDS_Data_Platform_ND-GAIN/data/nd-gain-joined.csv"))

```


Round value column for two numbers after the decimal

```{r}
nd_gain_data$value <- round(nd_gain_data$value, digits = 2)
```


**Challenge**

The dataset is slightly flawed because if a value is not measured one year, it simply gets carried over from the previous year

So simply think of filtering for the year 2020 as using the most recent year of available data (warning: this can lead to 20 year old data being visualized)


```{r}
nd_gain_most_recent <- nd_gain_data %>% 
  filter(year == 2020)
```

Shift dataframe so that each value in the indicator column becomes it's own column 
 - This gives us 190 rows (one country per row) with 48 indicators

```{r}
nd_gain_most_recent_wider <- nd_gain_most_recent %>% 
  pivot_wider(names_from = indicator,
              values_from = value)
```

Now our data is ready to visualize!

Here are a few interesting plots created by our SPICE 2023 Summer Students funded by the Pacific Intelligence and Innovation Initiative (P3I)


1. Histogram

This code chunk below creates a histogram that shows the distribution of dependency on external resource for health services across 190 countries colored by whether the country is Small Island Developing State (SIDS) or not

```{r}
ggplot(data = nd_gain_most_recent_wider, aes(x = dependency_on_external_resource_for_health_services,
                                     fill = sids_context
                                     )) +
  geom_histogram(bins = 65, color = "black") +
  theme_minimal()
```



2. Scatter Plots

Create a scatter plot with education on the x axis and the human development index on the y axis

```{r}
ggplot(data = nd_gain_most_recent_wider, aes(x = education,
                                     y = human_development_index)) +
  geom_point() +
  labs(title = "Education and Human Development Index, ND-GAIN Most Recent Data",
       x = "education",
       y = "human development index") +
  theme_minimal()


```
Now add a trend line with geom_smooth() and a p value and correlation coefficient with stat_cor()


```{r}
ggplot(data = nd_gain_most_recent_wider, aes(x = education,
                                     y = human_development_index)) +
  geom_point() +
  geom_smooth() +
  stat_cor(method = "pearson", output.type = "text", label.sep='\n') +
  labs(title = "Education and Human Development Index, ND-GAIN Most Recent Data",
       x = "education",
       y = "human development index") +
  theme_minimal()
```

Now split the plot into two for Small Island Developing States (SIDS) and non-SIDS 
 - Warning -> Be cautious of using trend lines, p-values, and correlation coefficients with <30 data points (in this case countries)

```{r}
ggplot(data = nd_gain_most_recent_wider, aes(x = education,
                                     y = human_development_index)) +
  geom_point() +
  geom_smooth() +
  stat_cor(method = "pearson", output.type = "text", label.sep='\n') +
  facet_wrap(~sids_context) +
  labs(title = "Education and Human Development Index, ND-GAIN Most Recent Data",
       x = "education",
       y = "human development index") +
  theme_minimal()
```
Now add color indicating GDP

```{r}
ggplot(data = nd_gain_most_recent_wider, aes(x = education,
                                     y = human_development_index,
                                     color = rural_population)) +
  geom_point() +
  geom_smooth() +
  stat_cor(method = "pearson", output.type = "text", label.sep='\n') +
  facet_wrap(~sids_context) +
  labs(title = "Education and Human Development Index, ND-GAIN Most Recent Data",
       x = "education",
       y = "human development index") +
  theme_minimal()
```

Lastly, make the plot interactive, adding label = Name so we can see which point is which country by hovering, and by naming it and inputting it into ggplotly()


```{r}
education_hdi_gdp_plot <- ggplot(data = nd_gain_most_recent_wider, aes(x = education,
                                     y = human_development_index,
                                     color = rural_population,
                                     label = Name)) +
  geom_point() +
  geom_smooth() +
  stat_cor(method = "pearson", output.type = "text", label.sep='\n', label.x = 20) +
  facet_wrap(~sids_context) +
  labs(title = "Education and Human Development Index, ND-GAIN Most Recent Data",
       x = "education",
       y = "human development index") +
  theme_minimal()

ggplotly(education_hdi_gdp_plot)
```


3. Interactive Scatter Plot without Trend Lines
- Sometimes, we don't intend to show a trend with a scatter plot


Use a scatter plot to display projected change of sea level rise and engagement in international environmental conventions, color by SIDS or non-SIDS


```{r}
engagement_impacts_plot <- ggplot(data = nd_gain_most_recent_wider, aes(x = engagement_in_international_environmental_conventions, 
                                             y = projected_change_of_sea_level_rise_impacts, 
                                             color = sids_context)) +
  geom_point() +
  labs(title = "Projected Change of Sea level Rise Impacts / Engagement in International Environmental Conventions",
       subtitle = "ND-GAIN Data",
       x = "engagement in international environmental conventions",
       y = "projected change of sea level rise impacts") +
  theme_minimal()

ggplotly(engagement_impacts_plot)
```



4. Subset Data and Create a Bar Chart

Subset Data for only SIDS

```{r}
nd_gain_most_recent_wider_sids <- nd_gain_most_recent_wider %>% 
  filter(sids_context == "sids")
```

Create Bar Chart showing Food Import Dependency in SIDS 

```{r}
ggplot(data = nd_gain_most_recent_wider_sids, aes(x = food_import_dependency, 
                                                  y = reorder(Name,food_import_dependency))) + 
  geom_bar(stat = "identity", fill = "steelblue", color = "black") +
  labs(title = "Food Import Dependency, SIDS",
       x = "proportion of cereal consumption obtained from imports",
       y = "")+
  theme_minimal() 
  
```





